---
title: "00_setup_data"
author: "Jagadeesh Puvvula"
date: "2026-05-31"
output: pdf_document
---

# Unzip Cell Ranger output and inspect the folder structure
# Each sample has filtered_feature_bc_matrix/ and raw_feature_bc_matrix/ subfolders
```{r}
zip_path  <- "~/Documents/snRNA_seq/cellranger_count_matrices.zip"
data_root <- "~/Documents/snRNA_seq/cellranger_data"

unzip(zip_path, exdir = data_root)

# List directories to confirm the path pattern before loading
list.dirs(data_root, recursive = TRUE)
```

# Locate the 20 filtered matrices and run per-sample QC
# Loading one sample at a time keeps memory use low on a local machine
```{r}
library(Seurat)
library(Matrix)
library(tidyverse)
library(scDblFinder)

data_root <- "~/Documents/snRNA_seq/cellranger_data"

# Find one filtered_feature_bc_matrix folder per sample; name each by its parent folder
sample_dirs <- list.dirs(data_root, recursive = TRUE) %>%
  str_subset("filtered_feature_bc_matrix$")
names(sample_dirs) <- basename(dirname(sample_dirs))

stopifnot(length(sample_dirs) == 20)   # sanity check: expect 20 samples
print(names(sample_dirs))

# process_one: load counts, build Seurat object, filter low-quality nuclei,
# tag sample/ancestry/fetal-sex metadata, and remove doublets per sample
obj_list <- imap(sample_dirs, process_one)
```

# Merge all 20 QC'd samples into a single object and checkpoint to disk
```{r}
merged <- merge(obj_list[[1]], obj_list[-1], add.cell.ids = names(obj_list))
rm(obj_list); gc()                                          # free memory
saveRDS(merged, "~/Documents/snRNA_seq/merged_qc.rds")      # save progress
```

# Normalize with SCTransform (variance-stabilizing transformation)
# Run sequentially (no parallel workers) to avoid exporting a huge object to each